Optimal forward modeling of gravitational fields using tesseroids with adaptive discretization

### 2014-09-02 body.tex
Corr to methods text and plot GLQ nodes in sample
Added some text to the part with the tesseroid equations. Added the
location of the GLQ nodes to the sample figure to make the case for
using the tesseroid dimensions instead node spacing.

Citing articles in text \citet{Asgharzadeh2007} and with parenthesis
\citep{Braitenberg2011}.

@@

\begin{equation}
    g_{\alpha}(r,\phi,\lambda) = G \rho
        \int\limits_{\lambda_1}^{\lambda_2}
        \int\limits_{\phi_1}^{\phi_2}
        \int\limits_{r_1}^{r_2}
        \frac{\Delta_{\alpha}}{\ell^3} \kappa dr' d\phi' d\lambda'
        \ \ \ \alpha \in \{x,y,z\},
    \label{eq:tessgrav}
\end{equation}

@@

\begin{eqnarray}
    I_{\alpha\beta} &=&
        \left(
            \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
            \frac{\delta_{\alpha\beta}}{\ell^3}
        \right) \kappa, \\
    \Delta_x &=& r' K_{\phi} , \\
    \Delta_y &=& r' \cos \phi' \sin(\lambda' - \lambda) , \\
    \Delta_z &=& r' \cos \psi - r, \\
    \ell &=& \sqrt{r'^2 + r^2 - 2 r' r \cos \psi} , \\
    \cos\psi &=& \sin\phi\sin\phi' + \cos\phi\cos\phi'
                 \cos(\lambda' - \lambda) , \\
    K_{\phi} &=& \cos\phi\sin\phi' - \sin\phi\cos\phi'
                 \cos(\lambda' - \lambda), \\
    \kappa &=& {r'}^2 \cos \phi'.
\end{eqnarray}

@@

Equations
\ref{eq:tesspot},
\ref{eq:tessgrav},
and
\ref{eq:tesstensor}
can be numerically integrated using
the 3D Gauss-Legendre Quadrature (GLQ)
\citep{Asgharzadeh2007}.
The GLQ
approximates the integral by
a weighted sum
\citep{Hildebrand1987},
e.g.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figs/vary-height-and-order}
    \caption{
        Example of the effect of varying
        the computation height
        and the number of nodes in the Gauss-Legendre Quadrature
        (i.e., the GLQ order).
        The $g_{xy}$ component was produced by a
        $7^\circ \times 7^\circ \times 20\ km$ tesseroid
        with $2.67\ g.cm^{-3}$ density
        and top at $z=0\ km$.
        The maps were calculated on a regular grid
        with $100\times100$ points.
        The GLQ order used was the same
        in the radial, longitudinal, and latitudial dimensions.
        a) Field calculated at $400\ km$ height using GLQ order 2.
        The effect is similar to that presented by \citet{Asgharzadeh2007}.
        b) At $150\ km$ height and order 2,
        the field resembles that of
        four point sources located at the GLQ nodes.
        This effect was shown by \citet{Ku1977}
        for the case of the $g_z$ component of a right rectangular prism.
        c) At $150\ km$ but with a higher GLQ order of 30,
        the field is similiar to that expected for a single mass source.
    }
    \label{fig:sample}
\end{figure*}

@@

We start by fixing the number of GLQ nodes to
$N^\lambda=N^\phi=N^r=2$.
In this way, the distances between adjacent nodes
are approximately the dimensions of the tesseroid
\citep[see][]{Wild-Pfeiffer2008}.
Thus, we may use the tesseroid dimensions
as proxies for the distance between nodes.
We define the dimensions of a tesseroid
in the radial, latitudinal, longitudinal directions, respectively,
as

>>>>

Using optimized Cartesian integration kernels by \citet{Grombein2013}.
Their advantage over the spherical integration kernels
\citep{Wild-Pfeiffer2008} is that they do not suffer from a singularity at the
poles.

@@

\begin{equation}
    g_{\alpha}(r,\phi,\lambda) = G \rho
        \int\limits_{\lambda_1}^{\lambda_2}
        \int\limits_{\phi_1}^{\phi_2}
        \int\limits_{r_1}^{r_2}
        \frac{\Delta\alpha}{\ell^3} \kappa dr' d\phi' d\lambda',
    \label{eq:tessgrav}
\end{equation}

@@

\begin{eqnarray}
    I_{\alpha\beta} &=&
        \left(
            \frac{3\Delta\alpha \Delta\beta}{\ell^5} -
            \frac{\delta_{\alpha\beta}}{\ell^3}
        \right) \kappa, \\
    \Delta x &=& r' K_{\phi} , \\
    \Delta y &=& r' \cos \phi' \sin(\lambda' - \lambda) , \\
    \Delta z &=& r' \cos \psi - r, \\
    \ell &=& \sqrt{r'^2 + r^2 - 2 r' r \cos \psi} , \\
    \cos\psi &=& \sin\phi\sin\phi' + \cos\phi\cos\phi'
                 \cos(\lambda' - \lambda) , \\
    K_{\phi} &=& \cos\phi\sin\phi' - \sin\phi\cos\phi'
                 \cos(\lambda' - \lambda), \\
    \kappa &=& {r'}^2 \cos \phi'.
\end{eqnarray}

@@

Equations
\ref{eq:tesspot},
\ref{eq:tessgrav},
and
\ref{eq:tesstensor}
must be numerically integrated
\citep{Grombein2013}.
We follow \citet{Asgharzadeh2007}
and use the 3D Gauss-Legendre Quadrature (GLQ) integration.
The GLQ
approximates the integral by
a weighted sum
\citep{Hildebrand1987},
e.g.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figs/vary-height-and-order}
    \caption{
        Example of the effect of varying
        the computation height
        and the number of nodes in the Gauss-Legendre Quadrature
        (i.e., the GLQ order).
        The $g_{xy}$ component was produced by a
        $7^\circ \times 7^\circ \times 20\ km$ tesseroid
        with $2.67\ g.cm^{-3}$ density
        and top at $z=0\ km$.
        The maps were calculated on a regular grid
        with $100\times100$ points.
        The GLQ order used was the same
        in the radial, longitudinal, and latitudinal dimensions.
        Black circles represent the horizontal location of the GLQ nodes
        (not shown for GLQ order 30).
        a) Field calculated at $400\ km$ height using GLQ order 2.
        The effect is similar to that presented by \citet{Asgharzadeh2007}.
        b) At $150\ km$ height and order 2,
        the field resembles that of
        four point sources located at the GLQ nodes.
        This effect was shown by \citet{Ku1977}
        for the case of the $g_z$ component of a right rectangular prism.
        c) At $150\ km$ but with a higher GLQ order of 30,
        the field is similar to that expected for a single mass source.
    }
    \label{fig:sample}
\end{figure*}

@@

We start by fixing the number of GLQ nodes to
$N^\lambda=N^\phi=N^r=2$.
In this way, the distances between adjacent nodes
can be roughly approximated by the dimensions of the tesseroid
(Fig.~\ref{fig:sample}a-b).
Thus, we may use the tesseroid dimensions
as proxies for the distance between nodes.
We define the dimensions of a tesseroid
in the radial, latitudinal, longitudinal directions, respectively,
as
